---
title: "Clustering  for Text Files"
author: "Gizem Güleli"
date: "2023-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Bonus TASK 

```{r Install and load packages ,warning=FALSE, results=FALSE, message=FALSE}

#install.packages(c("tm", "slam", "cluster", "wordcloud"))


# Load necessary libraries
library(tm)
library(slam)
library(cluster)
library(wordcloud)

```


```{r Read the text files}

# Get the list of text files
files <- list.files("text_files", pattern = "\\.txt", full.names = TRUE)

# Read the content of each text file with explicit encoding
documents <- sapply(files, function(file) {
  lines <- readLines(file, warn = FALSE, encoding = "UTF-8")
  iconv(lines, to = "UTF-8", sub = "byte")
})

```

```{r Preprocessing on Data to analyze}

# Create a corpus
corpus <- Corpus(VectorSource(documents))

# Cleaning
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removeWords, stopwords("english"))
corpus <- tm_map(corpus, stripWhitespace)

```
```{r}

# Create a document-term matrix
dtm <- DocumentTermMatrix(corpus)

# Convert the document-term matrix to a matrix
dtm_matrix <- as.matrix(dtm)

# compute distance between document vectors
distance <- dist(dtm_matrix)


```


# Analyze the Dataset

```{r - Hierarchical Clustering Method }

set.seed(123)

# Hierarchical clustering using Ward's method
hc <- hclust(distance, method = "ward.D")

# Cut the dendrogram to get clusters
hclustering <- cutree(hc, 2)

# Print the results

cat("Hierarchical Clustering Results:\n")
for (i in 1:2) {
  cluster_files <- files[hclustering == i]
  cat("Cluster", i, "files:\n")
  for (file in cluster_files) {
    cat(" -", file, "\n")
  }
  cat("\n")
}


```

```{r K-means Clustering}

set.seed(123)

# K-means clustering
kclustering <- kmeans(distance, centers = 2, nstart = 100)

# Print the results
cat("\nK-means Clustering Results:\n")
for (i in 1:2) {
  cluster_files <- files[kclustering$cluster == i]
  cat("Cluster", i, "files:\n")
  for (file in cluster_files) {
    cat(" -", file, "\n")
  }
  cat("\n")
}


```

```{r}
# Install and load the required libraries
install.packages(c("ggplot2", "dplyr"))
library(ggplot2)
library(dplyr)

# Read the content of each text file with explicit encoding
documents <- lapply(files, function(file) {
  lines <- readLines(file, warn = FALSE, encoding = "UTF-8")
  iconv(lines, to = "UTF-8", sub = "byte")
})

# Create a dataframe with file names and hierarchical clustering results
hierarchical_df <- data.frame(File = files, Cluster = cutree(hc, 2))

# Create a dataframe with file names and k-means clustering results
kmeans_df <- data.frame(File = files, Cluster = kclustering$cluster)

# Merge the two dataframes on the "File" column
combined_df <- merge(hierarchical_df, kmeans_df, by = "File", all = TRUE)

# Print the combined dataframe
cat("Combined Clustering Results:\n")
print(combined_df)

```
I create the data.frame with result to make better comparison. However i recognized that both clustering method named the clusters oppositely. Thus i changed the numbers for k-means clustering.


```{r}
# Swap values in the "Cluster.y" column
combined_df$Cluster.y <- ifelse(combined_df$Cluster.y == 1, 2, 1)

# Print the updated combined dataframe
cat("Updated Combined Clustering Results:\n")
print(combined_df)
```
```{r}
# Find rows where Cluster.x and Cluster.y do not match
mismatched_rows <- combined_df[combined_df$Cluster.x != combined_df$Cluster.y, ]

# Print the mismatched rows
print(mismatched_rows)
```
The majority of files are clustered consistently between Hierarchical and KMeans Clustering. However five files (listed above ) are identified as mismatched between the Hierarchical Clustering and K-means Clustering results. 

The mismatched files indicate differences in how the two clustering methods assigned these particular documents to clusters. The files may have characteristics that make them ambiguous or challenging for clustering algorithms to group consistently.

```{r}

# Print the counts for each cluster
cat("\nCount of files in each cluster:\n")
for (i in 1:2) {
  cat("Cluster", i, "files in Hierarchical Clustering:", sum(combined_df$Cluster.x == i), "\n")
  cat("Cluster", i, "files in KMeans Clustering:", sum(combined_df$Cluster.y == i), "\n\n")
}


```
```{r}

# Summary table of the top 5 words for each cluster

p_words <- colSums(dtm_matrix) / sum(dtm_matrix)

cluster_words <- lapply(unique(hclustering), function(x){
  rows <- dtm_matrix[hclustering == x, ]
  
  # For memory's sake, drop all words that don't appear in the cluster
  rows <- rows[, colSums(rows) > 0]
  
  colSums(rows) / sum(rows) - p_words[colnames(rows)]
})

cluster_summary <- data.frame(
  cluster = unique(hclustering),
  size = as.numeric(table(hclustering)),
  top_words = sapply(cluster_words, function(d){
    paste(
      names(d)[order(d, decreasing = TRUE)][1:10], 
      collapse = ", "
    )
  }),
  stringsAsFactors = FALSE
)

print("Cluster Summary:")
print(cluster_summary)

# Wordcloud for each cluster
for (i in unique(hclustering)) {
  cat("Wordcloud for Cluster", i, ":\n")
  wordcloud(
    words = names(cluster_words[[i]]), 
    freq = cluster_words[[i]], 
    max.words = 10, 
    rot.per = 0.2, 
    colors = brewer.pal(6, "Dark2")
  )
  cat("\n")
}

```

Cluster 1 is  Political Topics:

Words like "will," "going," "people," and "country" suggest discussions about the future, the nation, and its people. Words like "want," "America," and "know" indicate expressions of desires, patriotism, and a quest for knowledge.

Cluster 2 is about Business and Management:

Words like "data," "example," and "management" indicate a focus on data and project management. Words like "however," "problem," and "point" suggest an emphasis on problem-solving and addressing challenges, possibly using specific tools or methods like "ibis."


