---
title: "Regular Expressions"
author: "Gizem Güleli"
date: "2023-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Regular expressions (RegEx) are powerful tools for pattern matching and text manipulation, especially in the context of web scraping. This document explores the application of RegEx in different scenarios.

For those looking to enhance their skills and experience with regular expressions in a fun and engaging manner, I highly recommend Regex Crossword. It offers interactive puzzles that can significantly boost your RegEx proficiency.



```{r,include=FALSE,warning=FALSE, results=FALSE, message=FALSE}
# Load the necessary packages
library(tidyverse)
library(stringr)
```


# 1. Exercise

A vector of strings is given, and we'll use the str_view() function to perform various RegEx operations

```{r}

vector <- c("emoticon", ":)", "symbol", "$^$")
writeLines((vector))


# Use the function str_view() and find in vector: 
# a) string of 3 characters with the letter o in the middle
str_view(vector, '.o.')

# b) expression "emoticon"
str_view(vector, "^emoticon$")

# c) expression ":)"
str_view(vector, "^\\:\\)$")

# d) expression "$^$"
str_view(vector, "^\\$\\^\\$$")


```


# 2. Exercise ----

A corpus of 980 words is given (stringr::words), and we'll use str_view() to find words based on specific RegEx patterns.

```{r}

# Use the function str_view() and find in the corpus:
# a) all words containing the expression "yes" (add the parameter match=T)
str_view(stringr::words, "yes")

# b) all words starting with "w"
str_view(stringr::words, "^w")

# c) all words ending with "x"
str_view(stringr::words, "x$")

```


# 3. Exercise ----

Another corpus of 980 words is given (stringr::words), and we'll use str_view() for more complex RegEx patterns.

```{r}


# Use the function str_view() and find in the corpus:
# a) all words starting with a vowel
str_view(stringr::words, '^[aeiouAEIOU]')

# b) all words that start only with a consonant
str_view(stringr::words, '^[^aeiouAEIOU]')

# c) all words ending with "ing" or "ise"
str_view(stringr::words, '(ing|ise)$')

# d) all words ending with "ed" but not with "eed"
str_view(stringr::words, '[^e]ed$')

# -------------------------------------------------#
```

These exercises showcase the versatility of regular expressions in extracting specific patterns from text data, a crucial skill in data manipulation and analysis.

By practicing these exercises and exploring resources like Regex Crossword, you'll gain a deeper understanding of regular expressions, enhancing your data manipulation and analysis capabilities.






